---
title: "Clinical EDA"
date: "October 30, 2018"
output: html_document
---
  
```{r}
source("setup.r")
```

## Let's read in the clinical data

```{r}
clinicalRaw = read_csv(pe.clinicalDataPath, col_types = cols())
nrow(clinicalRaw)
head(clinicalRaw)
```

### Filter Clinical

49 of the mothers were excluded due to lack of adequate maternal sample. Exclude these from clinical


```{r}
clinical = pe.getClinicalInfo()
skim_tee(clinical)
```


### Now let's create a numerical version of the clinical data for PCA and other methods


```{r}
clinicalNumeric = pe.getClinicalAsNumerics()
clinicalNumeric[1:5,]
```


### Let's see any correlations in the clinical data

In the plot below, we see only expected correlations.

- Preeclampsia negatively correlates with gestAge and babyWgtGram
- If it's your first baby (primaGravida) you are likely younger (momAge)
- If your baby has IUGR it weighs less (babyWgtGram)
- IUGR and preeclampsia are positively correlated
- A longer gestAge correlates with a heavier birthweight (babyWgtGram)



```{r}
corrplot(cor(clinicalNumeric), method="circle", type="upper")
```


### We see correlation between ethnicities, as we would expect.

Let's check correlation of patients just as a sanity check. We should not see anything too extreme.

```{r}
# todo: try different row/column removal strategies
corrplot(pe.getNumericAlleleMatrix() %>% t %>% cor, method="circle", type="upper")
```



