---
title: "Maternal Cardiovascular-Related Single Nucleotide Polymorphisms, Genes, and Gene Regulatory Networks Associated with Early-Onset Preeclampsia"
output: html_notebook
---

# Maternal Cardiovascular-Related Single Nucleotide Polymorphisms, Genes, and Gene Regulatory Networks Associated with Early-Onset Preeclampsia

## Preeclampsia review

- Early signs of preeclampsia include high blood pressure and protein in urine
- Caused by the placenta
- 2 - 8% worldwide
- 63k deaths worldwide
- Earlier occurrence associated with worse outcomes
- 50 women per year die in US out of 4M births

## Early-Onset Preeclampsia (EOPE)
- A subset of PE
- Distinguishing characteristics:
-- Develops before 34 weeks of gestation. (Late Onset is after 37 weeks)
-- EOPE is often complicated by IUGR (Interuterine Growth Restriction)
-- EOPE has clear placental pathology
-- Leision of uteralplacental malperfusion

## Idea

- EOPE may have an underlying genetic predisposition, (studies show family history a factor).
- Many previous genetic investigations have looked at genes involved in ~10 processes such as angiogenesis and lipid metabolism
- There are some known associations between preeclampsia and cardiovascular and metabolic disease such as chronic hypertension and diabetes, and women with a history of preeclampsia more likely to get these diseases
- There are racial differences in rates of preeclampsia, NHOPI and Filipinos have increased risk

## Observations

- 60 mothers, first gestation, who delivered full-term normal birthweight infants
- Cases: 31 women with EOPE
 - Ethnically diverse
  - 15 Filipino
  - 6 White
  - 4 other Asian
  - 3 Native Hawaiian and other Pacific Islanders
- Controls: 29 controls

## Data collected

- MetaboChip with 196,7250 SNPs associated with metabolic and cardiovascular traits
 - All raw data is in an Illumina FinalReport txt file
   - 400MB uncompressed
   - TSV format (skip the first 9 lines)
   - 11.8M lines. 1 line for each SNP.
- Clinical data in clinical.csv

## Goal

- Find associations with EOPE
  - In SNPS (cardiovascular and metabolic in particular)
  - Genes
  - Regulatory pathways

## Results

- No SNPs found to be correlated (after corrections)
- Through gene-based tests:


## Sections


```{r child = 'manhattan-plots.rmd'}
```

```{r child = 'pcas.rmd'}
```

```{r child = 'qqplot.rmd'}
```

```{r child = 'gwas-tests.rmd'}
```

```{r child = 'clinical-eda.rmd'}
```

```{r child = 'phylotree.rmd'}
```

```{r child = 'tsne.rmd'}
```

```{r child = 'snp-eda.rmd'}
```

```{r child = 'metabochip-eda.rmd'}
```

```{r child = 'xgboost.rmd'}
```

```{r}
# rmarkdown::render('main.rmd')
```

